---
title: "Bayesian ANN-HMM"
author: "Al-Ahmadgaid B. Asaad"
date: "23 April 2016"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
## Introduction
Chapter Title: <font color = "#FFA700"><b>Bayesian Inference and Basic Definitions</b></font><br/>
Thesis Title: **BAYESIAN INFERENCE OF ARTIFICIAL NEURAL NETWORKS AND HIDDEN MARKOV MODELS HYBRIDS**<br/><br/>
by **Al-Ahmadgaid B. Asaad** (`alstatr.blogspot.com`; `alasaadstat@gmail.com`)<br/>
<br/><br/>
This notebook is meant to provide the source code for the Bayesian Inference of MCMC methods in Chapter 3.

## Monte Carlo Simulation
The following codes simulate the monte carlo method by approximating the area under the Gaussian distribution between -1.96 and 1.96.

```{r monte carlo, fig.height=5, fig.width=5, echo = FALSE}
set.seed(123)

library(magrittr)
library(grid)
library(gridExtra)
library(mvtnorm)
library(lattice)
library(plot3D)
library(KernSmooth)

draws <- seq(1000, 100000, by = 100)
area <- draws %>% length %>% numeric

for (i in 1:(draws %>% length)) {
    samples <- rnorm(n = draws[i])
    area[i] <- sum(((samples > -1.96) & (samples < 1.96)) / (samples %>% length))
}

xyplot(area ~ draws, type = c("g", "l"), col = "black") %>%
  update(xlab = "Draws", ylab = "Area")
```

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

