---
title: <font color = "#645452">Chapter 3</font>
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
by **Al-Ahmadgaid B. Asaad** (`alstatr.blogspot.com`; `alasaadstat@gmail.com`; `alstated@gmail.com`). This notebook contains source codes used in the thesis.

| <code>Chapter Title</code> | <font color = "#FFA700">Bayesian Inference and Basic Definitions </font>                                         |
| ------------- |-----------------------------------------------------------------------------------|
| <code>**Thesis Title**</code>  | <font color = "#FFA700">**Bayesian Inference of Artificial Neural Networks and Hidden Markov Models**</font> |



## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
library(animation)

saveGIF({
  for(i in 1:100){
    curve(sin(x), from = -5 + (i * 0.05), to = 5 + (i * 0.05), col = "red", ylab = "")
    curve(cos(x), from = -5 + (i * 0.05), to = 5 + (i * 0.05), add = TRUE, col = "blue", ylab = "")
    legend("topright", legend = c("sin(x)", "cos(x)"), fill = c("red", "blue"), bty = "n")
  }
}, interval = 0.1, ani.width = 550, ani.height = 350)
```


## Slide with Plot

```{r pressure}
plot(pressure)
```

